---
title: Map Functions (Purrr) and Letters
author: ''
date: '2020-02-29'
slug: map-functions-purrr-and-letters
categories: []
tags: []
---

```{r include=FALSE}
library(tidyverse)
library(gt)
```



For my problem set in Gov 1005, I learned how to interate over a tibble (table dataframe) using map functions from purrr. Ordinarily, I would have used a for loop or even worked with lists, but this opened my eyes to how succinct and useful map_* functions are!

The first thing I did was create a few function that I would reference later. 

The first function get_letters() takes an argument "a" that samples "a" letters from the alphabet and then using the base min() function to return the smallest letter (closest to A, farthest from Z). I didn't know this but there is a built-in vector "letters" in r that has all 26 letters already preassigned inside of it. Really helpful! 

The second function roll_alpha_dice() takes two arguments. It calls get_letters() "n" times, passing through the number of letters you initially want to sample "a". This uses the map_chr() function that returns a character vector storing n letters. 

```{r}


get_letters = function(a = 3){
  min(sample(letters, a))
}

roll_alpha_dice = function(n, a = 3){
  map_chr(1:n, ~get_letters(a))
}

```



Next up, I am going to use these functions to simulate a game. Let's say that there are four players (A,B,C,D). For each player, I will get 100 hundred letters and store them in a tibble. In this same tibble declaration, I can use another map function, map_lgl(), to iterate over the entire length of A's column and assign a logical (TRUE or FALSE) to a column called "vowel" if, unsurprisingly, that letter is a vowel. As always, we can use %in% to check to see if . (a row in the column A) is a in the string vector l.   
```{r}
n = 100
l = c("a","e","i","o","u")
x <- tibble(A=roll_alpha_dice(n), B=roll_alpha_dice(n), 
  C = roll_alpha_dice(n), D = roll_alpha_dice(n), vowel = map_lgl(A, ~. %in% l))

x %>% 
  head(5) %>% 
  gt()

```

```{r include=FALSE}
s = sum(x$vowel)

```


Turns out that there were `r s` vowels A's column.  



The coolest part is coming up. Now, I will make a bigger tibble of length 10,000. The same players are playing.
Okay, so what if I wanted to iterate over the entire dataframe and check how many times every single player got a vowel on the same roll? Can I do that? Of course I can! Using map!


So, the map_lgl() iterates over the length of the dataframe, a vector containing each row is returned--the "." is the index of the row. The base function all() checks if all letters in the vector are in l (is a vowel), and then I turn the result into a table that counts the true and false instances. 

```{r}

#I didn't need to name the columns but just for visual

n = 10000
x <- tibble(A=roll_alpha_dice(n),B=roll_alpha_dice(n),
    C = roll_alpha_dice(n), D = roll_alpha_dice(n))

y = prop.table(table(
  map_lgl(1:n, ~all(x[., ]  %in% l ))))[2]

```


Using this process, I simulated that the probability that they all roll a vowel is `r y * 100` %. 

Wicked right! 

I feel like this usually when a programmer breaks out her for loops, but nope, she doesn't even need them.
